---
title: "Working with Dates and Regex"
author: "Colleen Mossor"
date: "April 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir="../")

library(lubridate)
library(dplyr)
library(ggplot2)
library(DT)

```

## Purpose
Explore Biketown data looking at the mean miles travelled per week during the summer of 2018 06/2018 through 09/2018. 

```{r pull data, eval = F, echo = F}
# multiple isses with this. set eval = F to have it not run
setwd("../")
source("code/fetch_biketown.R")
get_data(start = "06/2018", end = "09/2018")
```

```{r pull all files, echo = F}
# for this function to work, must set the working drive to the same location
# also needs full file path

setwd("C:/Users/cocaldwell/Desktop/R Training/plotting/data/biketown/")
folder <- "C:/Users/cocaldwell/Desktop/R Training/plotting/data/biketown/"
# ^assigning the name "folder" to be used in the next section

filenames <-  list.files(path = folder, pattern = "*.csv", all.files=F,
                         full.names = F, recursive = F, ignore.case = F)
# ^assigns "filenames" as a list of files with a csv extension

read_csv_filename <- function(filenames){
  ret <- read.csv(filenames, stringsAsFactors = F, strip.white = T,
                  na.strings = "")
  ret$Source <- filenames
  ret
}
# ^ custom function that pulls the lsit of file names created above, and read them

bike_raw <- plyr::ldply(filenames, read_csv_filename)
# ^combines files into a single dataframe to work with


# IF THIS HADN'T WORKED, PULL IN EACH SEPARATELY
# file1 <- read.csv ("file path")
# file2 <- read.csv ("file path")
# 
# bike_raw_new <-  bind_rows(file1, file2)

```

```{r data wrangling, include = F}
str(bike_raw)

# combine start date and start time into one column, same with end
# don't want to mess with the raw data frame, create a new object
bike_df1  <- bike_raw %>% 
  mutate(start.datetime = paste(StartDate, StartTime, sep = " "),
         end.datetime = paste(EndDate, EndTime, sep = " "))
  # ^ paste vs paste0, paste0 has no separaters, paste does

str(bike_df1)
# ^new columns show up as chr, need to change to date
# make sure to include timezone

# check acceptable timezone values
OlsonNames()

bike_df1$start.datetime <-  mdy_hm(bike_df1$start.datetime, tz ="America/Los_Angeles")
bike_df1$end.datetime <-  mdy_hm(bike_df1$end.datetime, tz ="America/Los_Angeles")  
str(bike_df1)

# duration is another date/time issue since it's in chr
bike_df1$Duration <- hms(bike_df1$Duration)
# ^throws error "Some strings failed to parse, or all strings are NAs" it winds up rouding the NA values to 1 minute which isn't good

# check for NAs in the raw data file
sum(is.na(bike_raw$Duration))
# ^there are 1088 NAs, but since you have start and end time you can calculate this


```

'lubridate' has three synonymous functions but they define very different actions:

1. duration()
    time space are in units of seconds, and there is no start date involved. For example converting        HMS to durtaion, you're not making a calculation.
2. interval()
    also measured in units of seconds, and provides the difference beteen start and end
2. period()
    measures time span in units larger than seconds, this is handy when accounting for daylight savings     time and leap year. 
    
```{r duration_calc}
# calc interval between start and end times
bike_df1$interval <-  interval(bike_df1$start.datetime, bike_df1$end.datetime)
head(bike_df1$interval)

# calc duration
bike_df1$duration.all <- as.duration(bike_df1$interval)
head(bike_df1$duration.all)

# calculate period
bike_df1$period <- as.period(bike_df1$duration.all)
head(bike_df1$period)

```

```{r floor_date}
# floor_date() is helpful for aggregating

bike_weekagg <- bike_df1 %>% 
  mutate(week.datetime = lubridate::floor_date(start.datetime, unit = "week"))
  # ^looks at all values and sets them to the first day of the week, sunday if first day of week
  # ^ceiling date will move it to the cloest date going forward
  # there's also a round date
head(bike_weekagg$week.datetime)

avg_weekly_dist <- bike_weekagg %>% 
  group_by(week.datetime) %>% 
  summarise(avg_dist = mean(Distance_Miles))
View(avg_weekly_dist)

avg_weekly_fig <- avg_weekly_dist %>% 
  ggplot(aes(x = as.Date(week.datetime), y = avg_dist))+
  # ^as.Date is important for the scale below
  geom_bar(stat = "identity", fill = "pink") +
  # ^must use stat=identity or you get a frequency
  scale_x_date(date_breaks = "1 week") +
  xlab("Week") + ylab("Average Distance") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(label=round(avg_dist,1)))
avg_weekly_fig 


```

